---
title: "purrr tutorial"
author: "Sehrish Kanwal"
date: "`r format(Sys.time(), '%a %Y-%b-%d')`"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This report documents the work I did to learn R's `purrr` package (https://jennybc.github.io/purrr-tutorial/index.html). Trying to tick-off one item from my to-do list :memo:

## Required packages

```{r}
library(purrr)
library(repurrrsive)
library(listviewer)
library(tibble)
library(jsonlite)
library(dplyr)
library(stringr)
library(lubridate)
#library(here)
```

`listviewer` is an interactive way to explore objects vs `str`. 
`wesanderson`, `got_chars` and `gh_users` are lists from library `repurrrsive`. 

# Exploring data

```{r}
str(wesanderson)
listviewer::jsonedit(wesanderson, mode = "view")

str(got_chars, list.len = 3)
listviewer::jsonedit(got_chars, mode = "view")

str(gh_users, max.level = 1)
listviewer::jsonedit(gh_users, mode = "view")
```

# map(): Extract elements

A template for basic map() usage is `map(YOUR_LIST, YOUR_FUNCTION)`

```{r}
# Extract elements with name "name" - restrict to first few elements
map(got_chars[1:4], "name")

# Extract elements on the basis of integer values
map(got_chars[5:8], 3)

# Using the pipe %>%
got_chars[1:4] %>% 
  map("name")
got_chars[5:8] %>% 
  map(3)
```

In summary, are two shortcuts for making the .f function that map() will apply:

  - provide “TEXT” to extract the element named “TEXT”
         equivalent to function(x) x[["TEXT"]]
  - provide i to extract the i-th element
         equivalent to function(x) x[[i]]

## Exercises

Use names() to inspect the names of the list elements associated with a single character. What is the index or position of the playedBy element? Use the character and position shortcuts to extract the playedBy elements for all characters.

```{r}
map(got_chars[1:4], "playedBy")
map(got_chars[1:4], 18)
```

What happens if you use the character shortcut with a string that does not appear in the lists’ names?

```{r}
map(got_chars[1:4], "namess")
```

## Type-specific map

map() always returns a list, even if all the elements have the same flavor and are of length one. But in that case, you might prefer a simpler object: an atomic vector.

```{r}
#returns a list
map(got_chars[9:12], "name")

#returns atomic vector
map_chr(got_chars[9:12], "name")
```

## Extract multiple values

```{r}
got_chars[[3]][c("name", "culture", "gender", "born")]

#using map
x <- map(got_chars, `[`, c("name", "culture", "gender", "born"))
str(x[16:17])

#using extract from magrittr
library(magrittr)
x <- map(got_chars, extract, c("name", "culture", "gender", "born"))
str(x[18:19])
```

## Data frame output

Above, we extrcated lists from `map` list. However, we can stack-up these results row-wise 

```{r}
map_df(got_chars, extract, c("name", "culture", "gender", "id", "born", "alive"))
```

Notice how the variables have been automatically type converted. It’s a beautiful thing. Until it’s not. When programming, it is safer, but more cumbersome, to explicitly specify type and build your data frame the usual way.

```{r}
got_chars %>% {
  tibble(
    name <- map_chr(., "name"),
    culture <- map_chr(., "culture"),
    gender <- map_chr(., "gender"),
    id <- map_int(., "id"),
    born <- map_chr(., "born"),
    alive <- map_lgl(., "alive")
  )
}
```

# Simplifying data

A great use of purrr’s map() functions is to dig information out of a non-rectangular data structure and create a neat data frame. 

## Name and position shortcuts

Who are these GitHub users?

We need to reach into each user’s list and pull out the element that holds the user’s name or, maybe, username. How?

```{r}
map(gh_users, "login")
```

# Specifying the function in map()

Let's now work on three more ways to specify general `.f` function in map():

  - an existing function
  - an anonymous function, defined on-the-fly, as usual
  - a formula: this is unique to purrr and provides a very concise way to define an anonymous function
  
In Game of Thrones character list, got_chars, each character can have aliases, which are stored in a vector in each character’s component. We pull out the aliases for three characters to use as our demo.

```{r}
aliases <- set_names(map(got_chars, "aliases"), map_chr(got_chars, "name"))
aliases <- aliases[c("Theon Greyjoy", "Asha Greyjoy", "Brienne of Tarth")]
```

## Existing function

Use a pre-existing function. Or, as here, define one ourselves, which gives a nice way to build-in our specification for the  collapse argument.

```{r}
my_fun <- function(x)
  paste(x, collapse = " | ")

map(aliases, my_fun)
```


## Anonymous function, conventional

Define an anonymous function on-the-fly, in the conventional way. Here we put our desired value for the collapse argument into the function defintion itself.

```{r}
map(aliases, function(x) paste(x, collapse = " | ")) 
```

## Anonymous function, formula

We saved possibly the best for last.

purrr provides a very concise way to define an anonymous function: as a formula. This should start with the ~ symbol and then look like a typical top-level expression, as you might write in a script. Use .x to refer to the input, i.e. an individual element of the primary vector or list.

```{r}
map(aliases, ~ paste(.x, collapse = " | "))
```

# List columns

Getting and preparing data. I stumbled onto an issue trying to read data using read_csv. I had to update R and re-install few packages.

```{r}
load("trump_tweets_df.rda")
#glimpse(trump_tweets_df)

tweets <- trump_tweets_df$text
tweets %>% head() %>% strtrim(70)

tweets <- tweets[c(1, 2, 5, 6, 198, 347, 919)]
tweets %>% strtrim(70)
```

