---
title: "BAM-featureCount"
author: "Sehrish Kanwal"
date: "`r format(Sys.time(), '%a %Y-%b-%d')`"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required packages

```{r}
library(purrr)
library(repurrrsive)
library(listviewer)
```

`listviewer` is an interactive way to explore objects vs `str`. 
`wesanderson`, `got_chars` and `gh_users` are lists from library `repurrrsive`. 

## Data

```{r}
str(wesanderson)
listviewer::jsonedit(wesanderson, mode = "view")

str(got_chars, list.len = 3)
listviewer::jsonedit(got_chars, mode = "view")

str(gh_users, max.level = 1)
listviewer::jsonedit(gh_users, mode = "view")
```

## map(): Extract elements

```{r}
# Extract elements with name "name" - restrict to first few elements
map(got_chars[1:4], "name")

# Extract elements on the basis of integer values
map(got_chars[5:8], 3)

# Using the pipe %>%
got_chars[1:4] %>% 
  map("name")
got_chars[5:8] %>% 
  map(3)
```

In summary, are two shortcuts for making the .f function that map() will apply:

  - provide “TEXT” to extract the element named “TEXT”
         equivalent to function(x) x[["TEXT"]]
  - provide i to extract the i-th element
         equivalent to function(x) x[[i]]










